\documentclass[../main/thesis_msc.tex]{subfiles}


\begin{document}

\chapter*{Acknowledgments}

During my effort for the completion of this Master Thesis many people have stand by me and contributed in one way or another. First and foremost, I would like to thank my supervisor Prof. Dr. Norbert Langer for accepting me as a member of his group, his guidance, and for always being there when he was needed. His teachings, encouragement, and trust during the last year were all invaluable and something that I will always hold dear.


It is difficult to express my gratitude towards my advisor, Dr. John Antoniadis, in just a few lines. Thank you for the leap of faith you took, by taking me under your wings as an inexperienced graduate student with no degree of intuition for the field whatsoever. Hopefully my mistakes, and constantly naive questions have not left a permanent mark on your tolerance levels.


I am eternally grateful to Priv.-Doz. Dr. J\"urgen Kerp for his trust and support by giving me the chance to improve my knowledge on radio-astronomical observations, and my communication skills as a teaching assistant. Our exciting conversations on topics not directly related to our work, opened my eyes to new perspectives. 

I want, also, to thank Prof. Dr. Michael Kramer for agreeing to be my thesis evaluator on such a short notice, Dr. G\"otz Gr\"afener for illuminating discussions on stellar winds, and David Ramon Aguilera Dena for sharing with me pieces of his code.


There are no words to express my thankfulness to my family. Without their support, nothing would be possible to begin with. I cannot but devote this thesis to them.


To my girlfriend Anastasia for putting up with my frequent meltdowns, for her understanding, and for being abroad and not with her all of those years, I give my love.


Finally, to Tilemachos, Thaleia, Kostas, and Eftychia thank you for making my everyday life more tolerable and fun.

\end{document}
